---
title: "1.4 — Utility Maximization — Class Content"
execute:
  freeze: auto

---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Monday, February 7, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 1](/assignments/01-problem-set) (on classes 1.1-1.4) is due by 11:59 PM Monday February 13 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we put all of our tools together thus far to solve the consumer's constrained optimization problem. To solve the problem the “traditional way” would be to use [Lagrangian multipliers](https://en.wikipedia.org/wiki/Lagrange_multiplier) and calculus and solve for the first order conditions. [Take the full Lagrangian equation and solve for all partial derivatives, set equal to 0.] We instead will find the solution by looking graphically, and use an algebraic rule that should make a lot of intuitive sense. That rule, equivalently expressed in two ways:

$$
\begin{align*}\frac{MU_x}{MU_y} &= \frac{p_x}{p_y}\\
\frac{MU_x}{p_x} &= \frac{MU_y}{p_y}\\ 
\end{align*}
$$
The first rule states that the slope of the indifference curve (left) equals the slope of the budget constraint (right). This is stating that the marginal benefit of consuming $x$ (left) equals the marginal cost (right) of consuming $y$, and that the individual is trading off $y$ for $x$ at the same rate as the market is.

The second (version of the) rule, solved for $x$ and $y$ terms on either side, shows that the marginal utility per dollar spent is equal across both (all) goods. 

Both forms are often called the **equimarginal rule**, that at the optimum choice, the marginal benefit equals the marginal cost (which is the marginal benefit of the next best foregone opportunity).

We will do some [practice](/practice/1.4-practice) to make sure you can solve this type of model, and then you are fully prepared for [Problem Set 1](/assignment/01-problem-set).

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 4.4-4.5 in Goolsbee, Levitt, and Syverson, 2019
- <i class="fas fa-external-link-square-alt"></i> [Math Review Guide: Nonlinear Functions & Optimization ](https://micros23.classes.ryansafner.com/resources/math/nonlinear.html)
- <i class="fas fa-external-link-square-alt"></i> [Math Review Guide: Exponents & Logarithms ](https://micros23.classes.ryansafner.com/resources/math/exponents.html)

## <i class="fas fa-dumbbell"></i> Practice

Today we will be working on [practice problems](/practice/1.4-practice). Answers will be posted later on that page.

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 1 Due Monday February 13
[Problem Set 1](/assignments/01-problem-set) (on classes 1.1-1.4) is due by 11:59 PM Monday February 13 on Blackboard Assignments.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Solving the Constrained Optimization Problem with Calculus](/resources/appendices/1.4-appendix.html#solving-the-constrained-optimization-problem-with-calculus)
- [General Case for $n$ Goods & Marshallian Demand Functions](/resources/appendices/1.4-appendix.html#general-case-for-n-goods-marshallian-demand)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.4-slides")
```

