---
title: "1.5 — Demand — Class Content"
execute:
  freeze: auto

---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Wednesday, February 9, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 1](/assignments/01-problem-set) (on classes 1.1-1.4) is due by 11:59 PM Monday February 13 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Solving a one-time constrained optimum problem might be useful in certain contexts, but the *dynamics* of change are far more interesting and useful: how do people optimally *respond* to changes in constraints? The remaining few classes before Exam 1 are about how people respond to changes in prices and income. 

An individual’s **demand function** (often just called **“demand”**) for a particular good describes how that person optimally responds to changes in market conditions. While in truth, many things factor into one's demand for a good, we are simplifying the model to relate a person's quantity demanded of a good to three things: the price of that good (“own” price), the price of *other* goods, and the person's income. When we look only at the relationship between quantity demanded and the good's price (holding constant the other two things — other prices, and income), this is the familiar **demand curve** you see on supply and demand graphs from principles.

Today, we look only at how changes in *income* or changes in the prices of *other* goods can affect your quantity demanded for a good.

In doing so, we also introduce a major empirical tool that we will return to many times: an *elasticity.* You've seen one elasticity before: the price elasticity of demand, and probably calculated it using the "midpoint formula." There's a lot more to elasticity (and a lot of other kinds of elasticities!) than that. We will see two varieties today, and will spend most of a lecture on price elasticity, two class periods from now.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 5.1, 5.2, 5.4 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-dumbbell"></i> Practice

Today we will be working on [practice problems](/practice/1.5-practice). Answers will be posted later on that page.


## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 1 Due Monday February 13

[Problem Set 1](/assignments/01-problem-set) (on classes 1.1-1.4) is due by 11:59 PM Monday February 13 on Blackboard Assignments.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.5-slides")
```

