---
title: "1.7 — Price Elasticity — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Dates
*Wednesday, February 15, 2023* and
*Monday, February 20, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 2](/assignments/02-problem-set) (on classes 1.5-1.7) is due by 11:59 PM Monday February 20 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we wrap up Unit 1 by talking more extensively about demand curves and simple demand functions (relating price and quantity demanded of a good, *ceterus paribus*). We also (re-)introduce the idea of **price elasticity of demand** that measures a person's responsiveness in consumption to a change in that good's price. However, unlike what you learned in principles^[Arc-price elasticity using the midpoint between two points on the demand curve: $$\cfrac{\frac{q_{2}-q_{1}}{\left(\frac{q_{1}+q_{2}}{2}\right)}}{\frac{p_{2}-p_{1}}{\left(\frac{p_{1}+p_{2}}{2}\right)}}$$
This actually calculates the elasticity of a single point, the one midway between $(q_1,p_1)$ and $(q_2,p_2)$.] we look at the elasticity of *any point on the demand curve* using a more useful formula:
$$\epsilon_{q,p}=\frac{1}{slope} \times \frac{p}{q}$$
where slope refers to the slope of the *inverse* demand curve (the one that we graph).

We will look at the key relationship between price elasticity of demand and total revenue, (Total expenditure, from the perspective of the buyer.) and work on some [practice problems](/practice/1.7-practice).

Next class will be a review session for [Exam 1](/assignments/01-exam), and you can mind more information about on the exam that page.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 2.2, 2.5, 5.5, in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 2 Due Monday February 20

[Problem Set 2](/assignments/02-problem-set) (on classes 1.5-1.7) is due by 11:59 PM Monday February 20 on Blackboard Assignments.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("1.7-slides")
```

