---
title: "2.2 — Production Technology — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Dates
*Wednesday, March 1, 2023 and Monday, March 6, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we continue with setting up the second stage of the firm's problem, the *cost-minimization problem*, where firms choose inputs $(l,k)$ that minimize total cost $(wl+rk)$ in order to produce the optimal amount of output $q^*$. For now, we are merely *assuming* the amount of output that is optimal, later we will explicitly solve for it as our first stage (*profit maximization problem*).

In doing so, we need to split up production activity into the *short-run*, where firms cannot change all inputs (we traditionally assume capital is fixed $\bar{k}$ and firms can only change labor $l)$, and the long run, where firms can change all inputs. We discuss short run production and associated input-based metrics: *marginal product* and *average product.*

In the long-run, firms can choose *many* combinations of inputs $(l,k)$ that yield a desired output. We introduce two tools that will remind you a lot like the tools from consumer theory: (1) an *isoquant curve*^[Call it an output curve to remember it better.] that, similar to an indifference curve, contains all the combinations of inputs that yield the same output, and (2) an *isocost line*^[Call it just a cost line to remember it better.] that, similar to a budget constraint, contains all the combinations of inputs that are the same total cost. We discuss some properties about each of these things and do some examples in anticipation of putting these tools together in the next class to solve the problem.

For some students, it takes this second time around with the tools of consumer theory for them to finally "click," and that's okay!

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 6.2-6.4 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Derivation of MRTS Equation (as ratio of marginal products)](/resources/appendices/2.2-appendix.html#derivation-of-mrts-equation-as-ratio-of-marginal-products)
- [Solving the Firm’s Problem in One Step](/resources/appendices/2.2-appendix.html#solving-the-firms-problem-in-one-step)
- [The Firm’s Supply and Demand Functions](/resources/appendices/2.2-appendix.html#the-firms-supply-and-demand-functions)
- [The Geometry of Short-Run Production](/resources/appendices/2.2-appendix.html#the-geometry-of-short-run-production)
- [The Relationship Between Marginal and Average](/resources/appendices/2.2-appendix.html#the-relationship-between-marginal-and-average)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.2-slides")
```

