---
title: "2.4 — Costs of Production — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Monday, March 20, 2023*
:::

::: callout-warning
# Upcoming Assignment
[Problem Set 3](/assignments/03-problem-set) (on classes 2.1-2.3) is due by 11:59 PM Wednesday March 22 on Blackboard Assignments.
:::


## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we cover costs before we put them together next class with revenues to solve the firm’s (first stage) *profit maximization problem*. While it seems we are adding quite a bit, and learning some new math problems with calculating costs, we will practice it more next class, when put together with revenues.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 7 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [From Total Product to Total Cost: Output Expansion Paths](/resources/appendices/2.4-appendix.html)
- [Marginal Cost and Variable Cost](/resources/appendices/2.4-appendix.html#marginal-cost-and-variable-cost)
- [The Relationship Between Returns to Scale and Costs](/resources/appendices/2.4-appendix.html#the-relationship-between-returns-to-scale-and-costs)
- [Cobb-Douglas Cost Functions](/resources/appendices/2.4-appendix.html#cobb-douglas-cost-functions)

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 3 Due Wednesday March 22

[Problem Set 3](/assignments/03-problem-set) (on classes 2.1-2.3) is due by 11:59 PM Wednesday March 22 on Blackboard Assignments.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.4-slides")
```

