---
title: "2.5 — Short Run Profit Maximization — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Wednesday, March 22, 2023*
:::

::: callout-warning
# Upcoming Assignments
[Problem Set 3](/assignments/03-problem-set) (on classes 2.1-2.3) is due by 11:59 PM Wednesday March 22 on Blackboard Assignments.

[Problem Set 4](/assignments/04-problem-set) (on classes 2.4-2.6) is due by 11:59 PM Wednesday March 29 on Blackboard Assignments.

:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we briefly discuss revenues before we put them together to solve the firm's *profit maximization problem*. We focus on the *short run*, where the firm incurs fixed costs and must stay in the market regardless. We derive the firm's (inverse) *supply curve*.

Next class will be about the long run, and finish up unit 2.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 8.1-8.3 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-dumbbell"></i> Practice

Today we will be working on [practice problems](/practice/2.5-practice). Answers will be posted later on that page.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Common Cost Assumptions](/resources/appendices/2.5-appendix.html#common-cost-assumptions)

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 3 Due Wednesday March 22

[Problem Set 3](/assignments/03-problem-set) (on classes 2.1-2.3) is due by 11:59 PM Wednesday March 22 on Blackboard Assignments.

### Problem Set 4 Due Wednesday March 29

[Problem Set 4](/assignments/04-problem-set) (on classes 2.4-2.6) is due by 11:59 PM Wednesday March 29 on Blackboard Assignments.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.5-slides")
```

