---
title: "2.6 — Long Run Industry Equilibrium — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Dates
*Monday, March 27 and Wednesday March 29.*
:::

::: callout-warning
# Upcoming Assignment

[Problem Set 4](/assignments/04-problem-set) (on classes 2.4-2.6) is due by 11:59 PM Wednesday March 29 on Blackboard Assignments.
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

We wrap up Unit 2 on Producers this week by bringing our *optimization* model of how firms maximize profits into the long run, when firms can enter or exit depending on profitability. We also now need to talk about the fact that our firm is not the *only* profit-maximizing firm in the market, so we derive an **equilibrium** model of the industry in the long run.

Famously, we see that in *competitive* industries, economic profits get driven to zero in the long run, as firms enter or exit any time there are profits or losses.

We also talk about the hard to understand, but extremely important, idea of economic rents.

Next class we will have a **review session** for [**Exam 2**](/assignments/02-exam).

**Update**: I have added the answers to the price elasticity of supply question [in the slides](https://microf22.classes.ryansafner.com/slides/2.6-slides.html#63) (that I had to skip for time):

<p style="text-align:center"><a target="_blank" href="/files/practice/2.6-practice-answers.html"><span class="btn btn-primary btn-lg"><i class="far fa-life-ring"></i> Price Elasticity of Supply Question Answers</a></span></p>

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 8.4-8.5 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-dumbbell"></i> Practice

Answers from last class' [practice problems](/practice/2.5-practice) are posted on that page.

## <i class="fas fa-sigma"></i> Appendix

See the online appendix for today's content:

- [Producer Surplus](/resources/appendices/2.6-appendix.html#producer-surplus)
- [What’s the Difference Between Producer Surplus and Profit?](/resources/appendices/2.6-appendix.html#whats-the-difference-between-producer-surplus-and-profit)
- [Economic Rents](/resources/appendices/2.6-appendix.html#economic-rents)
- [The Shape of the Long Run Industry Supply Curve: External Economies](/resources/appendices/2.6-appendix.html#the-shape-of-the-long-run-industry-supply-curve-external-economies)

## <i class="fas fa-laptop-code"></i> Assignments

### Problem Set 4 Due Wednesday March 29

[Problem Set 4](/assignments/04-problem-set) (on classes 2.4-2.6) is due by 11:59 PM Wednesday March 29 on Blackboard Assignments.

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("2.6-slides")
```

