---
title: "3.1 — The Supply and Demand Model — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance=minimal}
# Meeting Date
*Wednesday, April 5, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we bring supply and demand together to talk about equilibrium, and begin to discuss how markets work in the real world. In doing so, we will run a demonstration in class using the [kiviq](kiviq.us) software ([see instructions here](/files/slides/using-kiviq.pdf)).

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 3 in Goolsbee, Levitt, and Syverson, 2019

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides. 

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("3.1-slides")
```

