---
title: "3.3 — The Static Benefits of Markets: Welfare Economics & Market Failure — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Wednesday, April 12, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we consider how markets work and why (and when) they are so efficient at allocating resources and solving social problems. We take a crack at defining efficiency in a few ways: [allocative](https://en.wikipedia.org/wiki/Allocative_efficiency), [productive](https://en.wikipedia.org/wiki/Productive_efficiency), [Pareto](https://en.wikipedia.org/wiki/Pareto_efficiency), and [Kaldor-Hicks](https://en.wikipedia.org/wiki/Kaldor%E2%80%93Hicks_efficiency) efficiency. We then discuss [Welfare Economics](https://en.wikipedia.org/wiki/Welfare_economics) and I summarize three conditions about when markets are good & efficient:
1. Markets are competitive
2. Markets reach equilibrium
3. Markets do not have externalities

This leads us into a discussion of [market failure](https://en.wikipedia.org/wiki/Market_failure), when markets *aren’t* efficient because something prevents the price system from working properly. We talk about several types of market failures: collective action problems, public goods, and externalities.

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fa-book"></i> Ch. 14.2-14.7, Ch. 16 in Goolsbee, Levitt, and Syverson, 2019

Your textbook does cover efficiency and welfare economics in Chapter 14, in the context of general equilibrium theory, which is a bit too advanced for us to spend sufficient time to cover. Thus, I would not recommend reading that chapter unless you are prepared to dive into new tools (Edgeworth diagrams, production possibilities frontiers) that I will not cover in this course.

Chapter 16, however, is a pretty accessible discussion of externalities and public goods.

### Optional/Recommended Reading

- [<i class="fas fa-file-pdf"></i> Safner, 2020, "'Public Good,' or 'Good for the Public?' Political Entrepreneurship and the Public Funding of Scientific Research"](https://ryansafner.com/research/)
- [<i class="fas fa-file-pdf"></i> Safner, 2016, "Institutional Entrepreneurship, Wikipedia, and the Opportunity of the Commons"](https://ryansafner.com/research/)
- [<i class="fas fa-podcast"></i> *Freakonomics Radio* ep. 115: "Parking is Hell"](https://freakonomics.com/podcast/parking-is-hell-a-new-freakonomics-radio-podcast/)
- [<i class="fab fa-youtube"></i> MR University: "Trading Pollution: How Pollution Permits Paradoxically Reduce Emissions"](https://www.youtube.com/watch?v=9tUb3MDrgEc) 
- [<i class="fab fa-youtube"></i> MR University: "Externalities and Incentives: The Economics of COVID (Vaccines)"](https://www.youtube.com/watch?v=2EFG6dvtQ6M)
- [<i class="fab fa-youtube"></i> *The Lorax* (Dr. Seuss: "How Bad Can I Be?"](https://www.youtube.com/watch?v=W_HUdf89hI8)
- [<i class="fas fa-podcast"></i> *NPR*: "The Nobel Laureate Who Figured Out How To Deal With Annoying People"](https://www.npr.org/sections/money/2013/09/10/218526472/the-nobel-prizewinner-who-figured-out-how-to-deal-with-annoying-people)
- [<i class="fab fa-youtube"></i> MR University: "Women in Economics: Elinor Ostrom"](https://learn.mru.org/women-economics-series-elinor-ostrom/) (solving tragedy of the commons)^[Elinor Ostrom was the first woman to win an Economics Nobel prize. She was famous for her work in studying how good rules can prevent a tragedy of the commons for all sorts of common resources such as fisheries, forests, water, and other environmental resources.]
- [<i class="fas fa-file-pdf"></i> Iannaconne, 1992, "Sacrifice and Stigma: Reducing Free-riding in Cults, Communes, and Other Collectives"](/readings/Iannaconne-1992.pdf)
- [<i class="fas fa-podcast"></i> *Econtalk*: Iannaconne on the Economics of Religion](https://www.econtalk.org/the-economics-of-religion/)

The following Wikipedia entries can also provide more background on today’s concepts (which are well known and standard among economists):

### On Efficiency & Welfare:

- [<i class="fab fa-wikipedia-w"></i> Allocative efficiency](https://en.wikipedia.org/wiki/Allocative_efficiency)
- [<i class="fab fa-wikipedia-w"></i> Productive efficiency](https://en.wikipedia.org/wiki/Productive_efficiency)
- [<i class="fab fa-wikipedia-w"></i> Pareto efficiency](https://en.wikipedia.org/wiki/Pareto_efficiency)
- [<i class="fab fa-wikipedia-w"></i> Kaldor-Hicks efficiency](https://en.wikipedia.org/wiki/Kaldor-Hicks_efficiency)
- [<i class="fab fa-wikipedia-w"></i> Eminent domain](https://en.wikipedia.org/wiki/Eminent_domain)
- [<i class="fab fa-wikipedia-w"></i> Welfare economics](https://en.wikipedia.org/wiki/Welfare_economics)

### On Market Failures (and their Solutions):

- [<i class="fab fa-wikipedia-w"></i> Market Failure](https://en.wikipedia.org/wiki/Market_failure)
- [<i class="fab fa-wikipedia-w"></i> Public good](https://en.wikipedia.org/wiki/Public_good_(economics))
- [<i class="fab fa-wikipedia-w"></i> Externality](https://en.wikipedia.org/wiki/Externality)
- [<i class="fab fa-wikipedia-w"></i> Tragedy of the Commons](https://en.wikipedia.org/wiki/Tragedy_of_the_Commons)
- [<i class="fab fa-wikipedia-w"></i> *The Logic of Collective Action*](https://en.wikipedia.org/wiki/The_Logic_of_Collective_Action)
- [<i class="fab fa-wikipedia-w"></i> Coase Theorem](https://en.wikipedia.org/wiki/Coase_theorem)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("3.3-slides")
```
