---
title: "4.3 — Pricing Strategies — Class Content"
execute:
  freeze: auto
---

```{r, echo = F}
source(here::here("slides", "slide-links.R"))
```

::: {.callout-note appearance="minimal"}
# Meeting Date

*Monday, May 1, 2023*
:::

## <i class="fas fa-info-circle fa-lg"></i> Overview

Today we look at how firms price their products when they have market power (that is to say, how most firms in reality price their products). 

## <i class="fas fa-book-reader fa-lg"></i> Readings

- <i class="fas fas fa-book"></i> Ch. 10 in Goolsbee, Levitt, and Syverson, 2019

### Recommended Reading

- [<i class="fas fa-podcast"></i> *NPR Planet Money* 416: “Why The Price of Coke Didn't Change For 70 years”](https://www.npr.org/2019/05/01/719213730/episode-416-why-the-price-of-coke-didnt-change-for-70-years)
- [<i class="fas fa-podcast"></i> *NPR Planet Money* 653: “The anti-store”](https://www.npr.org/sections/money/2015/09/25/443519599/episode-653-the-anti-store)
- [<i class="fas fa-podcast"></i> *NPR Planet Money* 633: “The Birth and Death of the Price Tag”](https://www.npr.org/sections/money/2015/06/17/415287577/episode-633-the-birth-and-death-of-the-price-tag)

- [<i class="fab fa-youtube"></i> Why Trains are so Expensive](https://www.youtube.com/watch?v=fwjwePe-HmA)

## <i class="fas fa-chalkboard-teacher"></i> Slides

Below, you can find the slides in two formats. Clicking the image will bring you to the html version of the slides in a new tab. The lower button will allow you to download a PDF version of the slides.

::: callout-tip
You can type <kbd>h</kbd> to see a special list of viewing options, and type <kbd>o</kbd> for an outline view of all the slides.
:::

I suggest printing the slides beforehand and using them to take additional notes in class (*not everything* is in the slides)!

```{r}
#| echo: false
#| output: asis

slide_links("4.3-slides")
```
